Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2018-11-18 mprop/mprop.tex
Importing the IRL toolkit

\documentclass{mprop}
\usepackage{graphicx}
\usepackage{amssymb,amsmath,amsthm,mathtools}
\usepackage{url}
\usepackage{bm}
\usepackage{bbm}
\usepackage{booktabs}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}

@@

Following previous work on GP IRL
\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
entropy IRL model \cite{ziebart2008maximum}, under which we have that
\[ P(a | s) \propto \exp(Q_r(s, a)), \]
where
\begin{equation}
  Q_r(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s, a, s')V_r(s'),
\end{equation}
and $V_r(s)$ is a ``soft'' version of the Bellman backup operator, which can be
obtained by repeatedly applying the following equation until convergence:
\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
The likelihood of the data can then be written down as
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
\begin{equation} \label{pDr}
  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right).
\end{equation}
However, a reward function learned by maximising this likelihood is not
transferable to new situations
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
reward structure in a way that would allow reward predictions for previously
unseen states.

@@

Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
states $\mathbf{r}$. While the previous paper to consider this IRL model
computed maximum likelihood estimates for $\bm\lambda$ and $\mathbf{u}$, and
made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
VI to approximate the full posterior distribution $\posterior$.
\emph{Variational inference} is an approximation technique for probability
densities \cite{blei2017variational}. Let $\approximation$ be our approximating
family of probability distributions for $\posterior$ with its own hyperparameter
vector $\bm\nu$. Then the job of VI algorithms is to optimise $\bm\nu$ in order
to minimise the \emph{Kullback-Leibler} (KL) divergence between the original
probability distribution and our approximation.  KL divergence (asymmetrically)
measures how different the two distributions are, and in this case can be
defined as \cite{blei2017variational}
\[ \begin{split}
    \DKL(\approximation || \posterior) &= \Eq [\log\approximation -
    \log\posterior ] \\
    &= \Eq[\log\approximation - \log\pfull] \\
    &+ \Eq[\log p(\mathcal{D}, \mathbf{X_u})].
  \end{split}
\]
The last term is both hard to compute and constant w.r.t. $\approximation$
\cite{blei2017variational}, so we can remove it from our optimisation objective.
The negation of what remains is often called the \emph{evidence lower bound}
(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
  be interpreted as Lebesgue integrals over the entire sample space.}
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}
% TODO: are they actually Lebesgue? What about notation?

@@

As mentioned in the introduction, most IRL algorithms assume that the reward
function can be represented as a linear combination of features. This assumption
originated in one of the earliest papers on the topic by Ng and Russell
\cite{DBLP:conf/icml/NgR00}, which introduced several linear programming
approaches to the problem. The authors also noticed that often
multiple reward functions can explain the same behaviour, and suggested
heuristics for reward functions that are ``far away'' from reward functions that
do not fit the data.

A few years later, Abbeel and Ng \cite{DBLP:conf/icml/PieterN04} developed an
algorithm for the same formulation of the problem with a guarantee to converge
quickly. Neu and Szepesv\'ari \cite{DBLP:conf/uai/NeuS07} identified a weakness
in Abbeel and Ng's approach: the algorithm requires features to be
``appropriately'' scaled, and optimal scaling may not be known. Instead, they
suggest a way to combine IRL with \emph{apprenticeship learning}, i.e., a
supervised learning task for optimal \emph{policy} recovery (whereas IRL focuses
on recovering the reward function).

@@

\begin{table}
  \centering
  \begin{tabular}{llllll}
    \toprule
    Authors, year & Inducing points & Hyperparameters & Complexity \\
    \midrule
    Titsias, 2009 \cite{DBLP:journals/jmlr/Titsias09} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hensman et al., 2013 \cite{DBLP:conf/uai/HensmanFL13} & fixed & variational & $\mathcal{O}(m^3)$ \\
    Gal et al., 2014 \cite{DBLP:conf/nips/GalWR14} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hoang et al., 2015 \cite{DBLP:conf/icml/HoangHL15} & fixed & fixed & $\mathcal{O}(m^3)$ \\
    Cheng and Boots, 2017 \cite{DBLP:conf/nips/ChengB17} & variational & variational & $\mathcal{O}(nm_\alpha + nm_\beta^2)$ \\
    Hensman et al., 2017 \cite{DBLP:journals/jmlr/HensmanDS17} & fixed & variational & $\mathcal{O}(nm)$ \\
    Peng et al., 2017 \cite{DBLP:conf/icml/PengZZQ17} & variational & variational & $\mathcal{O}(m^3)$ \\
    \bottomrule
  \end{tabular}
  \caption{Summary of relevant VI approximations to GPs. For both inducing
    points and hyperparameters, 'fixed' means 'chosen before the algorithm
    starts' and 'variational' means 'included amongst the variational
    parameters'. Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of
    bases used to represent the GP's mean and covariance, respectively.}
  \label{table:approximations}
\end{table}

Table \ref{table:approximations} further summarises some of the recent and/or
influential GP approximation approaches that might be suitable for our GP IRL
model. In order to derive a reliable ELBO, the inducing points should be either
fixed or modelled by a probability distribution (none of the approaches do
that), while hyperparameters cannot be variational (the reason will become clear
in Section \ref{sec:proposed_approach}). We would also like to avoid having the
hyperparameters fixed, as learning them efficiently would introduce a separate
problem. As the approach by Hoang et al. does not seem to be easily extendable
to modelled hyperparameters, it is unsuitable to our needs.

@@

\section{Proposed Approach} \label{sec:proposed_approach}

%state how you propose to solve the software development problem. Show that your
%proposed approach is feasible, but identify any risks.

In order to properly investigate the difference between variational inference
and maximum likelihood estimation for the model, we keep other parts of the
model the same. Namely, we set the covariance function to a version of the
automatic relevance detection kernel
\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_{\bm\lambda}(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
where $\lambda_0$ is the overall ``scale'' factor for how similar or distant the
states are,
\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) \]
is a diagonal matrix that determines how relevant each feature is (where $d$
denotes the number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Our vector of hyperparameters for the
covariance function is then $\bm\lambda = (\lambda_0, \dots,
\lambda_d)^\intercal$. Similarly, we keep the expression for the prior of
$\bm\lambda$:
\begin{equation} \label{eq:theta_prior}
  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}

Next, we can rewrite the posterior by using the chain rule and Bayes' theorem in
order to get a better sense of what we are trying to approximate:
\[
  \begin{split}
    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u}
    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
    \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times
    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
    \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda,
    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) \times p(\bm\lambda |
    \mathbf{X_u}) \times p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
    \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \times p(\mathcal{D} | r)
    \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
  \end{split}
\]
Note that now there are only two unknown probability distributions:
$p(\mathcal{D} | \bm\lambda, \mathbf{X_u})$ and $p(\mathcal{D} | \bm\lambda,
\mathbf{X_u}, \mathbf{u})$, which can be computed as follows:
\begin{align*}
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
\end{align*}
This suggests the following form for the approximation:
\begin{equation} \label{eq:approximation}
  \approximation = q(\bm\lambda) \times q(\mathbf{u} | \bm\lambda) \times q(\mathbf{r}
  | \bm\lambda, \mathbf{u}).
\end{equation}

\subsection{Structure of the Approximating Distribution}

@@

\subsection{Evidence Lower Bound}

@@

  variables the expectation is taken over. Also note that throughout the derivation
  equality is taken to mean ``equality up to an additive constant''.}

@@

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient \cite{blei2017variational}. Unfortunately, some of the
terms in $\mathcal{L}$ are still left as expected values. \emph{Black box
  variational inference} (BBVI) \cite{DBLP:conf/aistats/RanganathGB14} suggests
a way to express the gradient as an expectation without having to take the
gradient of the posterior:
\[ \nabla_{\bm\nu} \mathcal{L} = \Eq[\nabla_{\bm\nu} \log \approximation
  (\log \pfull - \log \approximation)]. \]
The gradient of the ELBO then has a unbiased estimate
\[ \nabla_{\bm\nu} \mathcal{L} \approx \frac{1}{S} \sum_{s=1}^S
  \nabla_{\bm\nu} \log \approximationS (\log \pfullS - \log
  \approximationS) \]
computed by drawing $S$ Monte Carlo samples $(\bm\lambda_s, \mathbf{u}_s,
\mathbf{r}_s) \sim \approximation$.

The main theorem behind the BBVI trick is the Lebesgue's dominated convergence
theorem \cite{royden1988real}, which---under some conditions---allows one to
claim that:
\[ \lim_{n \to \infty} \int_E f_n = \int_E \lim_{n \to \infty} f_n. \]
We can use the same idea on parts of $\mathcal{L}$ that we cannot take the
derivative of otherwise. In this case, we want a gradient instead of a limit,
and the integral represents an expected value.

@@

\begin{enumerate}
\item Prove all applications of the dominated convergence theorem.
\item Implement the algorithm as a more specialised version of Algorithm 2 in
  the black box VI paper.
\item Evaluation
\item Experiment with various improvements mentioned in the black box VI paper,
  other VI papers, especially the ones about GPs.
\item Write/finalise the paper.
\end{enumerate}

\subsection{Evaluation}

@@

The algorithms will be compared on variations of two commonly used fictional
scenarios: object world \cite{DBLP:conf/nips/LevinePK11} and highway driving
behaviour \cite{DBLP:conf/icml/PieterN04,DBLP:conf/nips/LevinePK10}. While more
recent papers (especially those using deep learning) often use the binary world
benchmark as well \cite{wulfmeier2015maximum,DBLP:conf/uai/JinDAS17}, our
approach is not able to construct new features, and thus cannot perform well on
such a task.

>>>>

\documentclass{mprop}
\usepackage{graphicx}
\usepackage{amssymb,amsmath,amsthm,mathtools}
\usepackage{url}
\usepackage{bm}
\usepackage{bbm}
\usepackage{booktabs}
\usepackage[multiple]{footmisc}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}

@@

Following previous work on GP IRL
\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
entropy IRL model \cite{ziebart2008maximum}, under which we have that
\[ P(a | s) \propto \exp(Q_r(s, a)), \]
where
\begin{equation}
  Q_r(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s, a, s')V_r(s'),
\end{equation}
and $V_r(s)$ is a `soft' version of the Bellman backup operator, which can be
obtained by repeatedly applying the following equation until convergence:
\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
The likelihood of the data can then be written down as
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
\begin{equation} \label{pDr}
  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right).
\end{equation}
However, a reward function learned by maximising this likelihood is not
transferable to new situations
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
reward structure in a way that would allow reward predictions for previously
unseen states.

@@

Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
states $\mathbf{r}$. While the previous paper to consider this IRL model
computed maximum likelihood estimates for $\bm\lambda$ and $\mathbf{u}$, and
made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
VI to approximate the full posterior distribution $\posterior$.
\emph{Variational inference} is an approximation technique for probability
densities \cite{blei2017variational}. Let $\approximation$ be our approximating
family of probability distributions for $\posterior$ with its own hyperparameter
vector $\bm\nu$. Then the job of VI algorithms is to optimise $\bm\nu$ in order
to minimise the \emph{Kullback-Leibler} (KL) divergence between the original
probability distribution and our approximation.  KL divergence (asymmetrically)
measures how different the two distributions are, and in this case can be
defined as \cite{blei2017variational}
\[ \begin{split}
    \DKL(\approximation || \posterior) &= \Eq [\log\approximation -
    \log\posterior ] \\
    &= \Eq[\log\approximation - \log\pfull] \\
    &+ \Eq[\log p(\mathcal{D}, \mathbf{X_u})].
  \end{split}
\]
The last term is both hard to compute and constant w.r.t. $\approximation$
\cite{blei2017variational}, so we can remove it from our optimisation objective.
The negation of what remains is often called the \emph{evidence lower bound}
(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
  be interpreted as definite integrals over the entire sample space.}
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}

@@

As mentioned in the introduction, most IRL algorithms assume that the reward
function can be represented as a linear combination of features. This assumption
originated in one of the earliest papers on the topic by Ng and Russell
\cite{DBLP:conf/icml/NgR00}, which introduced several linear programming
approaches to the problem. The authors also noticed that often
multiple reward functions can explain the same behaviour, and suggested
heuristics for reward functions that are `far away' from reward functions that
do not fit the data.

A few years later, Abbeel and Ng \cite{DBLP:conf/icml/PieterN04} developed an
algorithm for the same formulation of the problem with a guarantee to converge
quickly. Neu and Szepesv\'ari \cite{DBLP:conf/uai/NeuS07} identified a weakness
in Abbeel and Ng's approach: the algorithm requires features to be
`appropriately' scaled, and optimal scaling may not be known. Instead, they
suggest a way to combine IRL with \emph{apprenticeship learning}, i.e., a
supervised learning task for optimal \emph{policy} recovery (whereas IRL focuses
on recovering the reward function).

@@

\begin{table}
  \centering
  \begin{tabular}{llllll}
    \toprule
    Authors, year & Inducing points & Hyperparameters & Complexity \\
    \midrule
    Titsias, 2009 \cite{DBLP:journals/jmlr/Titsias09} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hensman et al., 2013 \cite{DBLP:conf/uai/HensmanFL13} & fixed & variational & $\mathcal{O}(m^3)$ \\
    Gal et al., 2014 \cite{DBLP:conf/nips/GalWR14} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hoang et al., 2015 \cite{DBLP:conf/icml/HoangHL15} & fixed & fixed & $\mathcal{O}(m^3)$ \\
    Cheng and Boots, 2017 \cite{DBLP:conf/nips/ChengB17} & variational & variational & $\mathcal{O}(nm_\alpha + nm_\beta^2)$ \\
    Hensman et al., 2017 \cite{DBLP:journals/jmlr/HensmanDS17} & fixed & variational & $\mathcal{O}(nm)$ \\
    Peng et al., 2017 \cite{DBLP:conf/icml/PengZZQ17} & variational & variational & $\mathcal{O}(m^3)$ \\
    \bottomrule
  \end{tabular}
  \caption{Summary of relevant VI approximations to GPs. For both inducing
    points and hyperparameters, `fixed' means `chosen before the algorithm
    starts' and `variational' means `included amongst the variational
    parameters'. Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of
    bases used to represent the GP's mean and covariance, respectively.}
  \label{table:approximations}
\end{table}

% TODO: the real reason is that we want to use the prior
% TODO: the reasoning about inducing points is not clear either
Table \ref{table:approximations} further summarises some of the recent and/or
influential GP approximation approaches that might be suitable for our GP IRL
model. In order to derive a reliable ELBO, the inducing points should be either
fixed or modelled by a probability distribution (none of the approaches do
that), while hyperparameters cannot be variational (the reason will become clear
in Section \ref{sec:proposed_approach}). We would also like to avoid having the
hyperparameters fixed, as learning them efficiently would introduce a separate
problem. As the approach by Hoang et al. does not seem to be easily extendable
to modelled hyperparameters, it is unsuitable to our needs.

@@

\section{Proposed Approach} \label{sec:proposed_approach}
%state how you propose to solve the software development problem. Show that your
%proposed approach is feasible, but identify any risks.

In this section we show a feasible way to apply VI to the problem. Section
\ref{sec:preliminaries} defines several previously-missing parts of the model
and provides intuition for how the approximating distribution could be
structured. Section \ref{sec:structure} then defines the approximating
distribution, where some choices are justified by a body of literature, and some
are novel. Next, Section \ref{sec:elbo} derives a simplified expression for the
ELBO. Finally, in Section \ref{vi_algs} we describe our strategy for
optimising the ELBO.

\subsection{Preliminaries} \label{sec:preliminaries}

In order to properly investigate the difference between variational inference
and maximum likelihood estimation for the model, we keep other parts of the
model the same. Namely, we set the covariance function to a version of the
automatic relevance detection kernel
\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_{\bm\lambda}(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
where $\lambda_0$ is the overall `scale' factor for how similar or distant the
states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
matrix that determines how relevant each feature is (where $d$ denotes the
number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Our vector of hyperparameters for the
covariance function is then $\bm\lambda = (\lambda_0, \dots,
\lambda_d)^\intercal$. Similarly, we keep the expression for the prior of
$\bm\lambda$:
\begin{equation} \label{eq:theta_prior}
  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}

Next, we can rewrite the posterior by using the chain rule and Bayes' theorem in
order to get a better sense of what we are trying to approximate:
\[
  \begin{split}
    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u}
    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
    \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times
    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
    \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda,
    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) \times p(\bm\lambda |
    \mathbf{X_u}) \times p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
    \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \times p(\mathcal{D} | r)
    \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
  \end{split}
\]
Note that now there are only two unknown probability distributions:
$p(\mathcal{D} | \bm\lambda, \mathbf{X_u})$ and $p(\mathcal{D} | \bm\lambda,
\mathbf{X_u}, \mathbf{u})$, which can be expressed as follows:
\begin{align*}
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
\end{align*}
This suggests the following form for the approximation:
\begin{equation} \label{eq:approximation}
  \approximation = q(\bm\lambda) \times q(\mathbf{u} | \bm\lambda) \times q(\mathbf{r}
  | \bm\lambda, \mathbf{u}).
\end{equation}

\subsection{Structure of the Approximating Distribution} \label{sec:structure}

@@

\subsection{Evidence Lower Bound} \label{sec:elbo}

@@

  variables the expectation is taken over. Also note that throughout the
  derivation equality is taken to mean `equality up to an additive constant'.}

@@

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient \cite{blei2017variational}. Unfortunately, some of the
terms in $\mathcal{L}$ are still left as expected values. \emph{Black box
  variational inference} (BBVI) \cite{DBLP:conf/aistats/RanganathGB14} suggests
a way to express the gradient as an expectation without having to take the
gradient of the posterior:
\[ \nabla_{\bm\nu} \mathcal{L} = \Eq[\nabla_{\bm\nu} \log \approximation
  (\log \pfull - \log \approximation)]. \]
The gradient of the ELBO then has a unbiased estimate
\[ \nabla_{\bm\nu} \mathcal{L} \approx \frac{1}{S} \sum_{s=1}^S
  \nabla_{\bm\nu} \log \approximationS (\log \pfullS - \log
  \approximationS), \]
computed by drawing $S$ Monte Carlo samples $(\bm\lambda_s, \mathbf{u}_s,
\mathbf{r}_s) \sim \approximation$.

The main theorem behind the BBVI trick is the Lebesgue's dominated convergence
theorem \cite{royden1988real}, which---under some conditions---allows one to
claim that:
\[ \lim_{n \to \infty} \int_E f_n\,d\mu = \int_E \lim_{n \to \infty}
  f_n\,d\mu. \]
We can use the same idea on parts of $\mathcal{L}$, derivatives of which cannot
be taken otherwise. In our case, we want a gradient instead of a limit, and the
integral represents an expected value.

% TODO: finish this
Let us briefly examine the options available in case the theorem is not
applicable to some of our expected values.
\[ \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] \quad \text{and} \quad \sum_{i=1}^d
  \mathbb{E}[\log(\lambda_i + 1)] \]
come from a chosen prior distribution for $\bm\lambda | \mathbf{X_u}$.
Therefore, if needed, the distribution can be replaced with an
easier-to-deal-with alternative.

% what if the parameters of the covariance function are variational?

%mTKuum, log|Kuu|, Tr[KuuS] => p(u | lambda, X_u)
%tTKruKuum, V_r(s), T(s, a, s')V_r(s) => p(D | r) (DTC?)

We can then take the BBVI algorithm as
the basis for ours, keeping in mind the numerous improvement possibilities
provided both by the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14} and many
others (e.g., the ones covered in Section \ref{literature}).

@@

The work plan can be summarised as follows:
\begin{enumerate}
\item Derive expressions for the derivatives of the ELBO w.r.t. the variational
  parameters. \emph{Deliverables}: full derivations that can be included as
  supplementary material of the final paper, including formal proofs for
  applications of the dominated convergence theorem. \emph{Date}: 15~December.
\item Implement the algorithm as a more specialised version of Algorithm~2 in
  the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14}. As most of the relevant
  IRL algorithms have been implemented in
  MATLAB\footnote{\url{http://graphics.stanford.edu/projects/gpirl/irl_toolkit.zip}}\footnote{\url{https://github.com/jinming99/DGP-IRL}},
  it makes sense to do the same in order to have accurate time-sensitive
  comparisons between the algorithms. \emph{Deliverables}: a working
  implementation that can be integrated into the IRL toolkit by Levine et al.
  \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
  code in the paper. \emph{Date}: 1~February.
\item Evaluate the algorithm by comparing it to multiple alternatives, as
  detailed in Section \ref{sec:evaluation}. \emph{Deliverables}: multiple plots
  in an SVG or PDF format, along with code that produces them. \emph{Date}:
  1~April.
\item Write/finalise the paper. \emph{Deliverable}: a 14-page paper.
  \emph{Date}: around 23~April, the deadline has not been updated yet.
\end{enumerate}

\subsection{Evaluation} \label{sec:evaluation}

@@

The algorithms will be compared on variations of two commonly used fictional
scenarios: object world \cite{DBLP:conf/nips/LevinePK11} and highway driving
behaviour \cite{DBLP:conf/icml/PieterN04,DBLP:conf/nips/LevinePK10}. While more
recent papers (especially those using deep learning) often use the binary world
benchmark as well \cite{DBLP:conf/uai/JinDAS17,wulfmeier2015maximum}, our
approach is not able to construct new features, and thus cannot perform well on
such a task.
